Laplacian envelope

### 2015-06-18 Notes/laplace.tex
written up shotened heuristics

The $i$'th row of this equation reads
%
\begin{equation}\label{eq:detarow}
  \frac{\delta\eta_{i+1}+\delta\eta_{i-1}}{2} - (s+1)\delta\eta_i = 0,
\end{equation}
%
with the following exceptions
%
\begin{equation}\label{eq:detabnduni}
\begin{aligned}
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_{M/2+1}+\delta\eta_{M/2-1}}{2} - (s+1)\delta\eta_{M/2} &= \frac{1}{s},\\
  \frac{\delta\eta_{M/2+2}+\delta\eta_{M/2}}{2} - (s+1)\delta\eta_{M/2+1} &= -\frac{1}{s},\\
  \frac{\delta\eta_{M-1}}{2} - \prn{s+\frac{1}{2}}\delta\eta_M &= 0.
\end{aligned}
\end{equation}
%
These equations are invariant under $\delta\eta_{M+1-i} \to -\delta\eta_i$, so these two must be equal.
Then we have
%
\begin{equation}\label{eq:detahalf}
\begin{aligned}
  \frac{\delta\eta_{i+1}+\delta\eta_{i-1}}{2} - (s+1)\delta\eta_i &= 0,
  \quad\text{for}\quad 1<i<\frac{M}{2}\\
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_{M/2-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{M/2} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%

The first of these equations is solved by
%
\begin{equation}\label{eq:detagensol}
  \delta\eta_i = B\alpha^{i-1} + C\alpha^{1-i},
  \qquad\text{where}\qquad
  \alpha^2 - 2(s+1)\alpha + 1 = 0,
\end{equation}
%
with $B$ and $C$ determined by the boundary conditions provided by the last two equations of \eqref{eq:detahalf}:
%
\begin{equation}\label{eq:unicoeffs}
  B = -\frac{2\alpha}{(1+\alpha)(\alpha^{M/2}+\alpha^{-M/2})s},
  \qquad
  C = -\frac{2}{(1+\alpha)(\alpha^{M/2}+\alpha^{-M/2})s}.
\end{equation}
%
This can be used to compute the Laplace transform
%
\begin{equation}\label{eq:unilaplace}
  A(s) = \frac{2\eta^w_1(s)}{M} 
   = \frac{2}{M}\prn{\frac{1}{s}+B+C} 
   = \frac{2(\alpha^{M/2}-1)^2}{Ms(\alpha^M+1)}.
\end{equation}
%

@@

The $\eta^w_i(s)$ take the same general form \eqref{eq:detagensol}, but with different coefficients due to the different boundary conditions
%
\begin{equation}\label{eq:detabndsticky}
\begin{aligned}
  \frac{\varepsilon\delta\eta_2}{2} - \prn{s+\frac{\varepsilon}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_{M/2-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{M/2} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%
These are solved by
%
\begin{equation}\label{eq:stickycoeffs}
\begin{aligned}
  B &= -\frac{ 2 \alpha^{M/2} (\alpha-(1-\varepsilon)) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s}, \\
  C &= -\frac{ 2 \alpha^{M/2} (1-(1-\varepsilon)\alpha) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s},
\end{aligned}
\end{equation}
%
which leads to the Laplace transform
%
\begin{equation}\label{eq:stickylaplace}
  \hat{A}(s) = \frac{2\varepsilon}{2+(M-2)\varepsilon} \, 
      \frac{ (\alpha^{M/2}-1)^2 - (1-\varepsilon)\alpha(\alpha^{M/2-1}-1)^2 }{ (\alpha^M + 1) - (1-\varepsilon)\alpha(\alpha^{M-2}+1) }.
\end{equation}
%

We can maximise this \wrt $\varepsilon$.
This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
For small $s \ll M^{-2}$, it is given by
%
\begin{equation}\label{eq:stickyeps}
  \varepsilon = \frac{ 2(M-1)\sqrt{s} }{\sqrt{M(M-2)}} \prn{ 1 + \sqrt{\frac{(M-2)^3s}{M}} + \CO(M^2s)},  
\end{equation}
%
and the Laplace transform is 
%
\begin{equation}\label{stickyenv}
  \hat{A}(s) = (M-1) \prn{ 1 - \sqrt{M(M-2)s} + \CO(M^2s)}.
\end{equation}
%

\section{Shortened serial model}\label{sec:serial_shorten}

@@

The $\eta^w_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
%
\begin{equation}\label{eq:detabndshorten}
\begin{aligned}
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_3 + \varepsilon\,\delta\eta_1}{2} - \prn{s+\frac{1+\varepsilon}{2}}\delta\eta_2 &= 0, \\
  \frac{\delta\eta_{M/2-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{M/2} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%
These are solved by
%

>>>>

The $i$'th row of this equation reads
%
\begin{equation}\label{eq:detarow}
  \frac{\delta\eta_{i+1}+\delta\eta_{i-1}}{2} - (s+1)\delta\eta_i = 0,
\end{equation}
%
with the following exceptions
%
\begin{equation}\label{eq:detabnduni}
\begin{aligned}
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_{\frac{M}{2}+1}+\delta\eta_{\frac{M}{2}-1}}{2} - (s+1)\delta\eta_{\frac{M}{2}} &= \frac{1}{s},\\
  \frac{\delta\eta_{\frac{M}{2}+2}+\delta\eta_{\frac{M}{2}}}{2} - (s+1)\delta\eta_{\frac{M}{2}+1} &= -\frac{1}{s},\\
  \frac{\delta\eta_{M-1}}{2} - \prn{s+\frac{1}{2}}\delta\eta_M &= 0.
\end{aligned}
\end{equation}
%
These equations are invariant under $\delta\eta_{M+1-i} \to -\delta\eta_i$, so these two must be equal.
Then we have
%
\begin{equation}\label{eq:detahalf}
\begin{aligned}
  \frac{\delta\eta_{i+1}+\delta\eta_{i-1}}{2} - (s+1)\delta\eta_i &= 0,
  \quad\text{for}\quad 1<i<\frac{M}{2}\\
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%

The first of these equations is solved by
%
\begin{equation}\label{eq:detagensol}
  \delta\eta_i = B\alpha^{i-1} + C\alpha^{1-i},
  \qquad\text{where}\qquad
  \alpha^2 - 2(s+1)\alpha + 1 = 0,
\end{equation}
%
with $B$ and $C$ determined by the boundary conditions provided by the last two equations of \eqref{eq:detahalf}:
%
\begin{equation}\label{eq:unicoeffs}
  B = -\frac{2\alpha}{(1+\alpha)(\alpha^{\frac{M}{2}}+\alpha^{-\frac{M}{2}})s},
  \qquad
  C = -\frac{2}{(1+\alpha)(\alpha^{\frac{M}{2}}+\alpha^{-\frac{M}{2}})s}.
\end{equation}
%
This can be used to compute the Laplace transform
%
\begin{equation}\label{eq:unilaplace}
  A(s) = \frac{2\eta^w_1(s)}{M} 
   = \frac{2}{M}\prn{\frac{1}{s}+B+C} 
   = \frac{2(\alpha^{\frac{M}{2}}-1)^2}{Ms(\alpha^M+1)}.
\end{equation}
%

@@

The $\eta^w_i(s)$ take the same general form \eqref{eq:detagensol}, but with different coefficients due to the different boundary conditions
%
\begin{equation}\label{eq:detabndsticky}
\begin{aligned}
  \frac{\varepsilon\delta\eta_2}{2} - \prn{s+\frac{\varepsilon}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%
These are solved by
%
\begin{equation}\label{eq:stickycoeffs}
\begin{aligned}
  B &= -\frac{ 2 \alpha^{\frac{M}{2}} (\alpha-(1-\varepsilon)) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s}, \\
  C &= -\frac{ 2 \alpha^{\frac{M}{2}} (1-(1-\varepsilon)\alpha) }{ (1+\alpha) [(1+\alpha^{M}) - (1-\varepsilon) \alpha (1+\alpha^{M-2}) ]s},
\end{aligned}
\end{equation}
%
which leads to the Laplace transform
%
\begin{equation}\label{eq:stickylaplace}
  \hat{A}(s) = \frac{2\varepsilon}{[2+(M-2)\varepsilon]s} \, 
      \frac{ (\alpha^{\frac{M}{2}}-1)^2 - (1-\varepsilon)\alpha(\alpha^{\frac{M-2}{2}}-1)^2 }{ (\alpha^M + 1) - (1-\varepsilon)\alpha(\alpha^{M-2}+1) }.
\end{equation}
%

We can maximise this \wrt $\varepsilon$.
This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=1$:
%
\begin{equation}\label{eq:stickyvalid}
\begin{aligned}
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
    M\alpha[(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^M+1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M-2}+1)]
     \\ &+ 2(\alpha^{\frac{M}{2}}-1)^2(\alpha^M+1) = 0,\\
    s = &\, \frac{(\alpha-1)^2}{2\alpha}.
\end{aligned}
\end{equation}
%
The ``sticky'' models are optimal at values of $s$ smaller than this.
This is not quite the same as the lower boundary of the region \eqref{eq:univalid}, so there is a gap between them where the uniform model with $M$ states is optimal.
The width of this gap is insignificant for larger $M$.

For small $s \ll M^{-2}$, it is given by
%
\begin{equation}\label{eq:stickyeps}
  \varepsilon = \frac{ 2(M-1)\sqrt{s} }{\sqrt{M(M-2)}} \prn{ 1 + \sqrt{\frac{(M-2)^3s}{M}} + \CO(M^2s)},  
\end{equation}
%
and the Laplace transform is 
%
\begin{equation}\label{eq:stickyenv}
  \hat{A}(s) = (M-1) \prn{ 1 - \sqrt{M(M-2)s} + \CO(M^2s)}.
\end{equation}
%

\subsection{Shortened serial model}\label{sec:serial_shorten}

@@

The $\eta^w_i(s)$ take the same general form \eqref{eq:detagensol} except for $i=1$, and with different coefficients due to the different boundary conditions
%
\begin{equation}\label{eq:detabndshorten}
\begin{aligned}
  \frac{\delta\eta_2}{2} - \prn{s+\frac{1}{2}}\delta\eta_1 &= 0, \\
  \frac{\delta\eta_3 + \varepsilon\,\delta\eta_1}{2} - \prn{s+\frac{1+\varepsilon}{2}}\delta\eta_2 &= 0, \\
  \frac{\delta\eta_{\frac{M}{2}-1}}{2} - \prn{s+\frac{3}{2}}\delta\eta_{\frac{M}{2}} &= \frac{1}{s}.
\end{aligned}
\end{equation}
%
These are solved by
%
%
\begin{equation}\label{eq:shortensol}
\begin{aligned}
  \delta\eta_i &= B\alpha^{i-2} + C\alpha^{2-i} \qquad
        \text{for} \quad 2 \leq i \leq \frac{M}{2}, \\
  \delta\eta_1 &= \frac{2\alpha^{\frac{M}{2}}}
         {[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
  B &= - \frac{2\alpha^{\frac{M}{2}}(\alpha^2-\alpha+1 + \varepsilon(\alpha-1))}
         {(1+\alpha)[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
  C &= - \frac{2\alpha^{\frac{M-2}{2}}(\alpha^2-\alpha+1 - \varepsilon\alpha(\alpha-1))}
         {(1+\alpha)[\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s}, \\
\end{aligned}
\end{equation}
%
Which gives the following Laplace transform
%
\begin{equation}\label{eq:shortenlaplace}
  \hat{A}(s) = \frac{2 [\varepsilon(\alpha^{\frac{M}{2}}-1)^2
          + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2] }
          { [M-2(1-\varepsilon)]
            [\varepsilon(\alpha^{M}+1) + (1-\varepsilon)(\alpha^2-\alpha+1)(\alpha^{M-2}+1)]s }.
\end{equation}
%
This interpolates between \eqref{eq:unilaplace} at $\varepsilon=1$ and the same expression at $M-2$ at $\varepsilon=0$.

We can maximise this \wrt $\varepsilon$.
This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=0\text{ or }1$:
%
\begin{equation}\label{eq:shortenvalid}
\begin{aligned}
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
    M(\alpha^2-\alpha+1)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M}+1)
      = 0,\\
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} \propto &\,
    (M-2)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^{M-2}+1)
      = 0,\\
    s = &\, \frac{(\alpha-1)^2}{2\alpha},
\end{aligned}
\end{equation}
%
with the $\varepsilon=1$ case providing the lower boundary and $\varepsilon=0$ providing the upper boundary.
If these were really the models that maximised $A(s)$, the region of validity for $M$ would not overlap with that for $M-2$.
In this case, however, there is an overlap.
This is related to the fact that the video shows that the optimal models do not6 shorten by one state at a time, but several outgoing transition probabilities are reduced at once.

\subsection{Heuristic envelope}\label{sec:heuristicenv}

We can obtain a good approximation to the envelope by combining the results \autoref{sec:serial_uniform} and \autoref{sec:serial_sticky}, ignoring the complications of \autoref{sec:serial_shorten}.
We will also ignore the small gap between the regions of validity, \eqref{eq:univalid} and \eqref{eq:stickyvalid}.

We will phrase it in terms of $\snrb(\tau)$, reintroducing $r$:
%
\begin{equation}\label{eq:envunits}
  \snrb(\tau) = \frac{ \sqrt{N} \hat{A}\prn{\frac{1}{r\tau}} }{ r\tau },
  \qquad
  \alpha(\tau) = \frac{ 1 + r\tau +\sqrt{1+2r\tau} }{ r\tau }.
\end{equation}
%
Then, combining \eqref{eq:binaryenv}, \eqref{eq:unienv} and \eqref{eq:stickylaplace}, we find
%
\begin{equation}\label{eq:henv}
  \snrb(\tau) = \left\{
\begin{aligned}
  & \frac{\sqrt{N}}{1+r\tau}, & 
  & r\tau \leq \frac{1}{2\sinh^2\frac{y_*}{2}} , \\
  & \sqrt{N}\prn{ \frac{ 4\sinh^2 \frac{y_*}{2} }{ y_* \cosh y_* } } \log\alpha, & 
  \frac{1}{2\sinh^2 \frac{y_*}{2}} \leq & r\tau \leq \frac{1}{2\sinh^2 \frac{y_*}{M}} , \\
  & \max_\varepsilon \brc{\frac{2\sqrt{N}\varepsilon}{2+(M-2)\varepsilon} \,
      \frac{ (\alpha^{\frac{M}{2}}-1)^2 - (1-\varepsilon)\alpha(\alpha^{\frac{M-2}{2}}-1)^2 }{ (\alpha^M + 1) - (1-\varepsilon)\alpha(\alpha^{M-2}+1) }}, &
  \frac{1}{2\sinh^2 \frac{y_*}{M}} \leq & r\tau .
\end{aligned}
\right.
\end{equation}
%
For $M \gg 1$, and $r\tau$ much greater than the lower bound of each region, we can approximate this as 
%
\begin{equation}\label{eq:henvapproc}
  \snrb(\tau) \approx \left\{
\begin{aligned}
  & \frac{\sqrt{N}}{1+r\tau}, &
  & r\tau \leq 0.73 , \\
  & \frac{ 1.1\sqrt{N}}{ \sqrt{r\tau} } , &
  0.73 \leq & r\tau \leq 0.22 M^2 , \\
  & \frac{ \sqrt{N} (M-1) }{ r\tau }, &\quad
  0.22 M^2 \leq &r\tau .
\end{aligned}
\right.
\end{equation}
%
This is plotted in \autoref{fig:heuristicenv}.

\begin{figure}[tb]
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=0.8\linewidth]{LenvHeuristic.svg}\\
  \caption{Heuristic envelope.
  Analytic expressions are only valid to the right of each region.}\label{fig:heuristicenv}
\end{figure}
